---
title: "MultiDimensionality Assessment"
output: html_notebook
---

### Number of multidimensional-scales:
1) 3 (mirrors FAMCAT Applied Cognition subdomains)
2) 5 (mirrors hypothesized HoPE Cognition subdomains)

###	Magnitude of scale intercorrelations
1) r’s ranging from 0 to 0.2 
2) r’s ranging from 0.2 to 0.5
3) r’s ranging from 0.5 to 0.7
4) r’s ranging from 0.7 to 0.9

### Bank Design:
1) Peaked Information (skewed and leptokurtic)
2) Flat/ideal Information

### IRT Scoring Methods:
MAP (need to use prior for pickles to pickles)


### CAT Administration:
1) Seq. Unidim SE Stopping (stop administration scale by scale)
2) Determinant of Fisher information matrix (MD-CAT) - chun’s psychometrika paper - item selection & termination

### Sample Size for Calibration:
1) N=300
2) N=500
3) N=750
4) N=1,000

### Sample Size for CAT:
1) Uniform across theta
2) N=1,000 distributed across -3 to 3 by 0.5 theta increments
3) Population that mimics HoPE patients (mildly to severely impaired population TBD)


### Item Parameters:
A) Discrimination: U[1.1, 2.8] - one param per dimension
B) Difficulty: U[-2,-0.67]; U[-0.67, 0.67]; U[0.67, 2] - <0.5 between thresholds


### Outcome Evaluation Criteria:
1) Number of items ‘saved’ = UD-CAT length - MD-CAT length 
2) Recovery of Item parameters, person parameters via CAT & between-item interscale correlation

```{r setup}
# install/load packages ####

  #install.packages("MASS")
  library(MASS)
  
  #install.packages("mirt")
  library(mirt)
  
  #install.packages("psych")
  library(psych)

# load IRT scripts ####

source("CalcInfo.R")
source("CalcRespProb.R")

# Script Graveyard ####
# install.packages("mirtCAT")
# library(mirtCAT)
```


```{r sim}

# Design Elements ####

  # Number of dimensions
  dimensions=c(3,5)

  # Number of items in each dimension
  ni_dim=50

  # Scale-scale intercorrelation magnitude
  intercorrmag=list(
              "0-0.2"=c("upper"=0.0, "lower"=0.2),
              "0.2-0.5"=c("upper"=0.2, "lower"=0.5),
              "0.5-0.7"=c("upper"=0.5, "lower"=0.7),
              "0.7-0.9"=c("upper"=0.7, "lower"=0.9))
  
  # Shape of bank information
  bankinfoshape=c("flat","leptokurtic","skewed")
  
  # Sample size of calibration data
  calbsamplesize=c(300,500,750,1000)
  
  # Create factorial list of design conditions
  design_conditions=expand.grid(
    "dimensions"=dimensions, 
    "intercorrmag"=names(intercorrmag), 
    "bankinfoshape"=bankinfoshape, 
    "calbsamplesize"=calbsamplesize)


# Simulate Item Pool ####
# itempool=t(replicate(100000,{
#     difficulties=c(-3,-1.5,1.5,3)
#     while(all(diff(difficulties)>0.5)){
#       difficulties=sort(sample(seq(-3,3,0.01),4))}
#     names(difficulties)=paste0("cb",1:length(difficulties))
#     item=c("a"=sample(seq(1.1,2.8,0.01),1),difficulties,"NCAT"=length(difficulties)+1)
#     info=sapply(seq(-3,3,0.1), CalcInfo, item)
#     c(item,
#       "maxinfo"=max(info),
#       "locmaxinfo"=unname((which.max(info)-31)/10),
#       "cb_sd"=sd(difficulties))
#     }))
#saveRDS(itempool, "itempool.RDS")
itempool=readRDS("itempool.RDS")

# Simulate Calibration Data ####
# use mapply to cycle through factorial design conditions
# sim=mapply(
#   function(numdimensions, cormag, bankinfoshape, calbss){
#       print(paste(numdimensions, cormag, bankinfoshape, calbss ))
#       
#       simulees=NA
#       iteration=1
#       while(length(simulees)==1){
#       print(paste("iteration:",iteration));
#       
#       #generate empty matrix for scale-scale correlations
#       scaleintercormat=matrix(1,nrow=numdimensions,ncol=numdimensions)
# 
#       #generate random scale-scale correlations from a given range
#       scaleintercorr=sample(
#                       seq(
#                         strsplit(as.character(cormag),"-")[[1]][1],
#                         strsplit(as.character(cormag),"-")[[1]][2],
#                         0.01),
#                       (numdimensions^2-numdimensions)/2)
#       
#       #populate MVN matrix
#       scaleintercormat[lower.tri(scaleintercormat)]=scaleintercorr
#       scaleintercormat=t(scaleintercormat)
#       scaleintercormat[lower.tri(scaleintercormat)]=scaleintercorr
# 
#       #sim data
#       simulees=try(mvrnorm(n=calbss,mu=rep(0,numdimensions), Sigma=scaleintercormat))
#       
#       iteration=iteration+1;}
#       
# 
# #new item bank simulation
# 
# thetaquad=round(seq(-3,3,0.1),1)
# 
# banksample=split(1:nrow(itempool),rep(1:10,nrow(itempool)/10))
# 
# items=lapply(banksample, function(banksample){
#   #select item from item pool (1 million items) to suit the information shape
#   bankinfoshapedensity=switch(as.character(bankinfoshape),
#                           "flat"=rep(1/61, times=length(thetaquad)),
#                           "leptokurtic"=dnorm(thetaquad,0,2)/sum(dnorm(thetaquad,0,2)),
#                           "skewed"=(dbeta(thetaquad/4+2.5/4,2.5,4)+2.5/4)/(sum(dbeta(thetaquad/4+2.5/4,2.5,4)+2.5/4)))
# 
#   
#   
#   
#   itemsused=c()
# 
#   bankinfo=rep(0,times=length(thetaquad))
#   names(bankinfo)=thetaquad
#   
#   items=itempool[banksample,]
#   iteminfo=apply(items,1,function(i){
#     sapply(thetaquad, function(t){CalcInfo(t,i)})})
# 
#   #iterate over 'best items'
#   
#   for(item in 1:ni_dim){
#       deviance=numeric(length=ncol(iteminfo))
#       bankiteminfo=iteminfo
#       for(candidateitem in 1:ncol(iteminfo)){
#         bankiteminfo[,candidateitem]=iteminfo[,candidateitem]+bankinfo
#         deviance[candidateitem]=
#           sum(abs(bankinfoshapedensity-bankiteminfo[,candidateitem]/sum(bankiteminfo[,candidateitem])))
#         }
#       bestitem=which.min(deviance[!(1:ncol(iteminfo)%in%itemsused)])
#       itemsused=c(itemsused,bestitem)
#       bankinfo=bankiteminfo[,bestitem]
#   }
#   list("items"=items[itemsused,], "bankinfo"=bankinfo,"design"=bankinfoshapedensity, "deviance"=deviance)
# })
# 
# items=items[order(sapply(items, function(x){x$deviance[length(x$deviance)]}),decreasing=FALSE)[1:numdimensions]]
# 
#       
#       
# 
#      calbdata=lapply(1:numdimensions, function(dim){
#                 sapply(simulees[,dim], function(theta){
#                   apply(items[[dim]][["items"]], 1, function(item){
#                     sample(1:5, 1, prob=CalcRespProb(theta,unlist(item[1:6])))
#                   })
#                 })
#               })
# 
#       list("items"=items, #"bankinfo"=bankinfo, "save"=save,
#            "simulees"=simulees, "calbdata"=calbdata)
#   }, 
# design_conditions$dimensions,
# design_conditions$intercorrmag,
# design_conditions$bankinfoshape,
# design_conditions$calbsamplesize,
# SIMPLIFY = FALSE
# )
# 
# saveRDS(sim, "sim.RDS")
sim=readRDS("sim.RDS")

  
```  
  
``` {r calb}  
  
# Calibration ####
# 
# sim=readRDS("sim.RDS")
# library(mirt)
# library(parallel)
# 
# 
# calb=mclapply(sim, mc.cores = 96, function(i){
#   numdim=length(i$calbdata)
# 
#   if(numdim==3){
#     dat=t(do.call(rbind, i$calbdata))
#     colnames(dat)=c(1:150)
#     model <- mirt.model('
#       F1 = 1-50
#       F2 = 51-100
#       F3 = 101-150
#       COV = F1*F2, F1*F3, F2*F3')
#     cmod <- mirt(dat, model)  
#     
#   }
#   
#   if(numdim==5){
#     dat=t(do.call(rbind, i$calbdata))
#     colnames(dat)=c(1:250)
#     model <- mirt.model('
#       F1 = 1-50
#       F2 = 51-100
#       F3 = 101-150
#       F4 = 151-200
#       F5 = 201-250
#       COV = F1*F2, F1*F3, F1*F4, F1*F5, F2*F3, F2*F4, F2*F5, F3*F4, F3*F5, F4*F5')
#     cmod <- mirt(dat, model)    
#   }
#   
#   cmod
#   
# })
# saveRDS(calb, "calb.RDS")
calb=readRDS("calb.RDS")

```

``` {r results}




# d= -b*a
# d/a=-b
# -d/a=b




#process
1)simulate data for various conditions (cycle through mapply & expand.grid)
2)fit confirmatory mirt model, using output (sampled) FA loadings (items assigned to highest factor-item loading)
3)cat sim



coef(calb[[1]])

```

``` {r ScriptGraveyard}

#Old item selection by bank info design shape script ####

#       # items=replicate(numdimensions, simplify=FALSE, {
#       #   #select item from item pool (1 million items) to suit the information shape
#       #   bankinfoshapepts=switch(as.character(bankinfoshape),
#       #                             "flat"=round(seq(-3,3,length.out=ni_dim),1),
#       #                             "leptokurtic"=round(rnorm(ni_dim,0,0.25),1),
#       #                             "skewed"=round(rbeta(ni_dim,2,5),1))
#       #   
#       #   if(bankinfoshape=="flat"){
#       #     bankinfoshapepts=sample(bankinfoshapepts,length(bankinfoshapepts))
#       #   } else {
#       #     bankinfoshapepts=as.numeric(unlist(
#       #       apply(data.frame(sort(table(bankinfoshapepts),decreasing=TRUE)),1,
#       #             function(x){rep(x[1],times=x[2])})))}    
#       # 
#       #   bankinfoshapepts=bankinfoshapepts-min(bankinfoshapepts)
#       #   bankinfoshapepts=bankinfoshapepts*(3/max(bankinfoshapepts))
#       #   bankinfoshapepts=round(bankinfoshapepts,1)
#       #   
#       #   #select item from item pool (1 million items) to suit the information shape
#       #     items_used=c()
#       # 
#       #     items=matrix(NA, ncol=ncol(itempool), nrow=length(bankinfoshapepts))
#       # 
#       #     for(index in 1:length(bankinfoshapepts)){
#       #         t=bankinfoshapepts[index]
#       #         item=which(itempool[,"locmaxinfo"]==t & !(1:nrow(itempool) %in% items_used))
#       #         item=item[which.max(itempool[item,"maxinfo"])]
#       #         items_used=c(items_used,item)
#       #         items[index,]=itempool[item,]        
#       # 
#       #     }
#       #     colnames(items)=colnames(itempool)
#       #     items
#       # })
#       #     
#       #     
#       # #items=split(items,sort(rep(1:numdimensions,ni_dim)))
#       # 
#       # bankinfo=lapply(1:numdimensions, function(dim){
#       #           bankinfo=rowSums(
#       #               apply(items[[dim]],1,function(i){
#       #                 sapply(seq(-3,3,0.1), function(t){CalcInfo(t,i)})}))
#       #           names(bankinfo)=seq(-3,3,0.1)
#       #           bankinfo
#       #           })





#Old 'bank info optimization.R' script ####
# 
# thetaquad=round(seq(-3,3,0.1),1)
# 
# banksample=split(1:nrow(itempool),rep(1:10,nrow(itempool)/10))
# 
# items=lapply(banksample, function(banksample){
#   #select item from item pool (1 million items) to suit the information shape
#   bankinfoshapedensity=switch(as.character(bankinfoshape),
#                           "flat"=rep(1/61, times=length(thetaquad)),
#                           "leptokurtic"=dnorm(thetaquad,0,2)/sum(dnorm(thetaquad,0,2)),
#                           "skewed"=(dbeta(thetaquad/4+2.5/4,2.5,4)+2.5/4)/(sum(dbeta(thetaquad/4+2.5/4,2.5,4)+2.5/4)))
# 
#   
#   
#   
#   itemsused=c()
# 
#   bankinfo=rep(0,times=length(thetaquad))
#   names(bankinfo)=thetaquad
#   
#   items=itempool[banksample,]
#   iteminfo=apply(items,1,function(i){
#     sapply(thetaquad, function(t){CalcInfo(t,i)})})
# 
#   #iterate over 'best items'
#   
#   for(item in 1:ni_dim){
#       deviance=numeric(length=ncol(iteminfo))
#       bankiteminfo=iteminfo
#       for(candidateitem in 1:ncol(iteminfo)){
#         bankiteminfo[,candidateitem]=iteminfo[,candidateitem]+bankinfo
#         deviance[candidateitem]=
#           sum(abs(bankinfoshapedensity-bankiteminfo[,candidateitem]/sum(bankiteminfo[,candidateitem])))
#         }
#       bestitem=which.min(deviance[!(1:ncol(iteminfo)%in%itemsused)])
#       itemsused=c(itemsused,bestitem)
#       bankinfo=bankiteminfo[,bestitem]
#   }
#   list("items"=items[itemsused,], "bankinfo"=bankinfo,"design"=bankinfoshapedensity, "deviance"=deviance)
# })
# 
# items=items[order(sapply(items, function(x){x$deviance[length(x$deviance)]}),decreasing=FALSE)[1:numdimensions]]
# 
# ####plotting for lab
# png("out.png", 800,400)
# par(mfrow = c(1,2))
# 
# plot(thetaquad, items$'1'$design, type="l", ylab="info density", main=paste(bankinfoshape, "Design Bank Info"))
# 
# 
# plot(thetaquad,items$'1'$bankinfo, ylim=c(0,60), type="l", ylab="info", main=paste(bankinfoshape,"Simulated Bank Info"))
# lines(thetaquad,items$'2'$bankinfo, lty=2)
# lines(thetaquad,items$'3'$bankinfo, lty=3)
# dev.off()
# 





```


 
 
